An Infinite de Bruijn Sequence

### 2016-11-29 infdebruijn.tex
Cycle joining section

\usepackage{epsfig} 
\usepackage{amsmath} 
\usepackage{amssymb} 
\usepackage{amsthm}  
\usepackage{hyperref}
\usepackage{tikz}
\usetikzlibrary{automata,arrows,decorations.pathmorphing,decorations.markings,positioning}
\usetikzlibrary{shapes}
%\usepackage{balance}

@@

\theoremstyle{definition}
\newtheorem{definition}{Definition}     
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{observation}[theorem]{Observation}
\newtheorem{example}[theorem]{Example}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{corrolary}[theorem]{Corrolary}

@@

\newcommand{\R}{{\mathbb{R}}}
\newcommand{\N}{{\mathbb{N}}}
\newcommand{\C}{{\mathbb{C}}}
\newcommand{\Z}{{\mathbb{Z}}}

@@

\title{On Combinatorial Games and de Bruijn Sequences}

\author{Yotam Svoray and %Gera Weiss\\
Department of Computer Science \\
Ben Gurion University of the Negev\\
Israel
}

@@

\begin{document}

>>>>

\usepackage{epsfig} 
\usepackage{amsmath} 
\usepackage{amssymb} 
\usepackage{amsthm}  
\usepackage{hyperref}
\usepackage{tikz}
\usetikzlibrary{automata,arrows,decorations.pathmorphing,decorations.markings,positioning}
\usetikzlibrary{shapes}
%\usepackage{balance}
\usepackage{scrextend}
\addtokomafont{labelinglabel}{\sffamily}

@@

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}     
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{observation}[theorem]{Observation}
\newtheorem{example}[theorem]{Example}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{corrolary}[theorem]{Corrolary}

%\newcommand{\RotateLeft}{rlft}
\DeclareMathOperator{\RotateLeft}{RotLft}

@@

\newcommand{\R}{{\mathbb{R}}}
\newcommand{\N}{{\mathbb{N}}}
\newcommand{\Z}{{\mathbb{Z}}}

@@

\title{An Infinite de Bruijn Sequence}

\author{}

@@

\begin{document}
\maketitle

\section{A cycle joining Construction}

\begin{definition}
	A word in $\N^*$ is called a \emph{key word} if it is bigger in right-to-left lexicographical order than all of its rotations.
\end{definition}

\begin{definition}[Cycle Construction]
For $n\in \N$, let $k_1,k_2,\dots$ be an enumeration in right-to-left lexicographic order of all the key words of length $n$ and let $z(k_i)$ be the number of leading zeros in $k_i$. Define $C_0=\T{0^n}$ and, for $i>0$, let $C_i$ be the sequence obtained from $C_{i-1}$ by adding 
	$$\T{\RotateLeft(k_i,z(k_i)+1),\RotateLeft(k_i,z(k_i)+2),\dots,\RotateLeft(k_i,z(k_i))}$$ 
after the word $(\sigma-1)w$ where $\sigma w=\RotateLeft(k_i,z(k_i))$. \footnote{This is always possible because  $(\sigma-1)w$ is a rotation of a key word that is smaller than $k_i$ in left-to-right lexicographic order, i.e., it came earlier in the enumeration.}  
\end{definition}

\begin{proposition}
For every $l>0$ there is some $i>0$ such that the first $l$ words in $C_i$ are also the first $l$ words in $C_j$ for any $j>i$.
\end{proposition}

\begin{definition}
Let $C_\infty$ be the infinite sequence whose prefixes are all prefixes of an infinite number of elements in $\{C_0,C_1,\dots\}$.
\end{definition}

\begin{proposition}
The sequence $C_\infty$ is a de Bruijn sequence of order $n$, i.e., it is a Hamiltonian path over the infinite de Bruijn graph of order $n$ whose vertexes are the words $\N^n$ and whose edges are $\{ \T{ \sigma w, w \sigma'} \colon \sigma, \sigma' \in \N, w \in \N^{n-1} \}$.
\end{proposition}

\begin{definition}[Last on Necklace]
Let $last(w)$ be true if and only if $w=\RotateLeft(k_i,z(k_i))$ for some $i$. i.e., if $w$ is the last member of its necklace (set of rotations) in the sequence. 
\end{definition}

\begin{definition}
Let $nxt\colon \N^n \to \N^n$ be defined by
$$nxt(\sigma w)=\begin{cases} 
w(\sigma+1) & \text{if }last((\sigma+1)w) \\
w0          & \text{if }last(\sigma w) \\
w\sigma     & \text{otherwise} 
\end{cases}$$
\end{definition}

\begin{proposition}
For every $w\in \N^n$, $nxt(w)$ is the word that follows $w$ in the sequence $C_\infty$.
\end{proposition}

\begin{definition}
Let $prv\colon \N^n \to \N^n$ be defined by
$$prv(w\sigma)=\begin{cases} 
(\sigma-1)w & \text{if }last(\sigma w) \\
\sigma'w   & \text{if $\sigma =0$  and $\sigma'$ is the maximal such that $last(\sigma' w)$} \\
\sigma w     & \text{otherwise} 
\end{cases}$$
\end{definition}

\begin{proposition}
For every $w\in \N^n\setminus \{0^n\}$, $prv(w)$ is the word that precedes $w$ in the sequence $C_\infty$.
\end{proposition}

\begin{proposition}
Both $prv$ and $nxt$ can be computed in $O(n^2)$ time and space.
\end{proposition}

\begin{proposition}[Prefer Maximum]
For every $\sigma \in \N$ and $w \in \N^{n-1}$, the word $\sigma w$ comes before the word $(\sigma+1) w$ in $C_\infty$.
\end{proposition}
\begin{proof}
Let $k$ be a key word  of the form $0^z(k)\hat{w}$. We will prove the proposition for every rotation of $k$:

\begin{description} 
\item [First Case:] If $w=0^{z_1}\hat{w}0^{z_2}$ such that $z_1+z_2=z(k)$ and $z_1 \neq 0$ then we have that $w$ precedes  $w_1=0^{z_1-1}\hat{w}0^{z_2}1$ by the definition of $nxt$ and that $w_1$ precedes $1 0^{z_1-1}\hat{w}0^{z_2}$ by the order that we added the words in this necklace.
\item [Second Case:] If $w=\sigma u 0^{z(k)}v$ such that $v\sigma u=\hat{w}$ we have that
\begin{itemize}
	\item $w$ precedes $v \sigma u 0^{z(k)}$ because the latter is last on its necklace.
	\item $v \sigma u 0^{z(k)}$ precedes $\sigma u 0^{z(k)+|v|}$ because...
	\item $\sigma u 0^{z(k)+|v|}$ precedes $0(\sigma+1) u 0^{z(k)+|v|-1}$ because...
	\item $0(\sigma+1) u 0^{z(k)+|v|-1}$ precedes the word $u_2 (\sigma+1) u_1$ such that $u_1u_2=u0^{z(k)+|v|+1}$ and $last(u_2 (\sigma+1) u_1)$, by the order of the enumeration of the words in this necklace.
	\item $u_2 (\sigma+1) u_1$ precedes $0(\sigma+1)u_1 0^{|u_2|-1}$ because...
	\item $0(\sigma+1)u_1 0^{|u_2|-1}$ precedes $0 \tau_2 \dots \tau_l (\sigma+1)u_1$ where $u_2=\tau_1\dots\tau_l \sigma u_1=v\sigma u 0^{z(k)}$ because... 
	\item $0 \tau_2 \dots \tau_l (\sigma+1)u_1$ precedes $\tau_2 \dots \tau_l (\sigma+1)u_1\tau_1$ because... 
	\item $\tau_2 \dots \tau_l (\sigma+1)u_1\tau_1$ precedes $(\sigma+1)u_1\tau_1 \tau_2 \dots \tau_l=(\sigma+1) u 0^{z(k)}v$ because... 
\end{itemize}
\end{description}
